Expert-In-The-Loop Causal Discovery: \\ Iterative Model Refinement Using Expert Knowledge

### 2025-02-10 paper/main.tex
udpates

\paragraph{Both $ X $ and $ Y $ are continuous: }
When both $ X $ and $ Y $ are continuous, Pearson's correlation test can be
used and Pearson's correlation coefficient can be used at the effect size. When
$ \bm{Z} = \emptyset $, the correlation coefficient is defined as: $ r_{X, Y} =
\mathrm{cov}(X, Y) / (\sigma_X \sigma_Y) $. When $ \bm{Z} \neq \emptyset $,
partial correlation coefficient can be used. This is estimated by fitting two
regression models $ E_X: X \sim \bm{Z} $ and $ E_Y: Y \sim \bm{Z} $,
calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $ R_Y = Y - E_Y(\bm{Y})
$, and then computing the correlation coefficient between the residuals: $r_{X,
Y \rvert \bm{Z}} = r_{R_X, R_Y}$

@@

Polyserial (when one is ordinal and the other is continuous) and Polychoric
(when both are ordinal) correlation have been used to estimate the covariance
matrix between them \citep{Poon1987}. Both methods make the assumption that the
observed ordinal variable is a result of thresholding a latent normally
distributed continuous variable. Under this assumption, the methods then try to
estimate the covariance matrix while maximizing the likelihood of the dataset.
Using the estimated covariance matrix, $ \Sigma $ we can perform the CI test
and compute the Pearson's correlation coefficient as the effect size. When 
$\bm{Z} = \emptyset $, $ r_{X, Y} = \Sigma_{XY} / (\sqrt{\Sigma_{XX} \Sigma_{YY}}) $, 
	and when $ \bm{Z} \ne \emptyset $,
	$ r_{X, Y \rvert \bm{Z}} = - \Sigma^{-1}_{XY}/ (\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}) $.

>>>>

\paragraph{Both $ X $ and $ Y $ are continuous: }
When both $ X $ and $ Y $ are continuous, we can use a (partial) correlation
test for CI testing and Pearson's correlation coefficient can be used at the
effect size. When $ \bm{Z} = \emptyset $, the correlation coefficient is
defined as: $ r_{X, Y} = \mathrm{cov}(X, Y) / (\sigma_X \sigma_Y) $. When $
\bm{Z} \neq \emptyset $, partial correlation coefficient can be used. This is
estimated by fitting two regression models $ E_X: X \sim \bm{Z} $ and $ E_Y: Y
\sim \bm{Z} $, calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $ R_Y =
Y - E_Y(\bm{Y}) $, and then computing the correlation coefficient between the
residuals: $r_{X, Y \rvert \bm{Z}} = r_{R_X, R_Y}$

@@

Polyserial (when one is ordinal and the other is continuous) and Polychoric
(when both are ordinal) correlation have been used to estimate the covariance
matrix between them \citep{Poon1987}. Both methods make the assumption that the
observed ordinal variable is a result of thresholding a latent normally
distributed continuous variable. Under this assumption, the methods then try to
estimate the covariance matrix while maximizing the likelihood of the dataset.
Using the estimated covariance matrix, $ \Sigma $ we can perform a correlation
test for CI and compute the Pearson's correlation coefficient as the effect
size. When 
$\bm{Z} = \emptyset $, $ r_{X, Y} = \Sigma_{XY} / (\sqrt{\Sigma_{XX} \Sigma_{YY}}) $, 
	and when $ \bm{Z} \ne \emptyset $,
	$ r_{X, Y \rvert \bm{Z}} = - \Sigma^{-1}_{XY}/ (\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}) $.
